---
title: "Sesame Street Report [Draft]"
author: 'Armelle, Sara, Ibrohim'
date: 'February 05, 2024'
output: pdf_document
---

```{r}
#| echo: false
#| include: false
#| label: front-matter
#| warning: false
# clean up & set default chunk options
rm(list = ls())
knitr::opts_chunk$set(echo = FALSE)

# packages
library(tidyverse) # for example
library(ggformula) # for example
library(tinytex)   # for example
library(GGally)
```

## 1. Project Description

The project involves an observational study designed to assess the impact of
"Sesame Street" viewership on children's learning outcomes, particularly in
letters, numbers, and forms. The study is in the planning stage, aiming to
analyze existing data collected from children across five different sites,
including variables such as viewing frequency, setting, and encouragement to
watch, alongside pretest scores of vocabulary maturity (Peabody Picture
Vocabulary Test). The objective is to determine the program's effectiveness in
educational content delivery and identify areas for improvement. Results will
be used to enhance "Sesame Street's" educational focus and effectiveness, as
per the client's request for an upcoming board meeting presentation.

## 1.1 Research Questions

**Question 1:** How does viewership of "Sesame Street" influence children's
learning outcomes, specifically in terms of letters, numbers, and forms
knowledge, as measured by the percent increase in post-test scores compared to
pre-test scores?

**Question 2:** What is the percent achievable gain (PAG) in learning outcomes
for children who watch "Sesame Street," considering the maximum possible score
improvement based on their initial knowledge level?

## 1.2 Variables

| Variable Type       | Variable Name          | Description                                                                                      | Measurement/Recording                                    |
|---------------------|------------------------|--------------------------------------------------------------------------------------------------|----------------------------------------------------------|
| Explanatory         | Viewing Frequency      | Frequency of "Sesame Street" viewership by children.                                             | Recorded as the number of episodes watched per week.     |
| Explanatory         | Viewing Setting        | Context in which "Sesame Street" is watched (e.g., at home, in school).                         | Categorical variable (Home, School, Other).              |
| Explanatory         | Encouragement to Watch | Level of encouragement from parents or teachers for children to watch "Sesame Street."           | Binary variable (Yes, No).                               |
| Response            | Percent Increase       | Difference in test scores before and after the exposure period, adjusted for scale differences.  | Calculated as percent_post_test - percent_pre_test.      |

#  2. Exploratory Data Analysis (EDA)

The Exploratory Data Analysis (EDA) of the Sesame Street study integrates
comprehensive assessment and insights into children's learning improvements
attributed to the program. Through examining pre- and post-test outcomes over a
six-month interval among children aged 3-5, the study aims to quantify the
educational impact of Sesame Street on early learning metrics such as literacy
and numeracy. It meticulously considers variables like demographics, viewing
habits, and educational encouragement while addressing analytical challenges
like variance disparities, confounders, and multicollinearity. Initial findings
indicate variable program effectiveness, suggesting a correlation between
viewing frequency and educational advancement. The planned next steps involve
sophisticated statistical techniques to refine these insights, setting a robust
foundation for conclusive analysis and program enhancement strategies.

## Import Data
```{r}
#| echo: false
#| include: false
#| label: front-matter
#| warning: false
just_scores <- read.csv('test_scores.csv')
subset_scores <- subset(just_scores, 
                        grepl('let', test_name) | 
                        grepl('form', test_name) | grepl('numb', test_name))

subset_scores <- subset_scores |> 
  mutate(which_test = case_when(grepl('let', test_name) ~ 'letters',
                                grepl('form', test_name) ~ 'forms',
                                grepl('numb', test_name) ~ 'numbers'))

Data <- read.csv('clean_df.csv')

Data <-  Data |> 
  mutate(encour = case_when(encour == 0 ~ 'No',
                            encour == 1 ~ 'Yes'))

# make categorical variables categorical 
Data$site <-  as.character(Data$site)
Data$sex <-  as.character(Data$sex)
Data$viewcat <-  as.character(Data$viewcat)
Data$setting <-  as.character(Data$setting)
```

```{r}
ggplot(Data, aes(x = viewcat, y = let_diff, fill = viewcat)) +
  geom_boxplot(show.legend=FALSE) +
  xlab("View Time (higher value=more)") + 
  ylab("Improvement") + 
  labs(title = "Letters") + 
  theme(axis.text=element_text(size=22),
        axis.title=element_text(size=30,face="bold"),
        title=element_text(size=25,face="bold"))

ggplot(Data, aes(x = viewcat, y = numb_diff, fill = viewcat)) +
  geom_boxplot(show.legend=FALSE) +
  xlab("View Time") + 
  ylab("Improvement") + 
  labs(title = "Numbers") + 
  theme(axis.text=element_text(size=22),
        axis.title=element_text(size=30,face="bold"),
        title=element_text(size=25,face="bold"))

ggplot(Data, aes(x = viewcat, y = form_diff, fill = viewcat)) +
  geom_boxplot(show.legend=FALSE) +
  xlab("View Time") + 
  ylab("Improvement") + 
  labs(title = "Forms") + 
  theme(axis.text=element_text(size=22),
        axis.title=element_text(size=30,face="bold"),
        title=element_text(size=25,face="bold"))
```

Comparing improvement (post-test - pre-test) based on encouragement (1 = was encouraged)
```{r}

ggplot(Data, aes(x = encour, y = let_diff, fill = encour)) +
  geom_boxplot(show.legend=FALSE) +
  xlab("Encouragement") + 
  ylab("Improvement") + 
  labs(title = "Letters") + 
  theme(axis.text=element_text(size=22),
        axis.title=element_text(size=30,face="bold"),
        title=element_text(size=25,face="bold"))

ggplot(Data, aes(x = encour, y = numb_diff, fill = encour)) +
  geom_boxplot(show.legend=FALSE) +
  xlab("Encouragement") + 
  ylab("Improvement") + 
  labs(title = "Numbers") + 
  theme(axis.text=element_text(size=22),
        axis.title=element_text(size=30,face="bold"),
        title=element_text(size=25,face="bold"))

ggplot(Data, aes(x = encour, y = form_diff, fill = encour)) +
  geom_boxplot(show.legend=FALSE) +
  xlab("Encouragement") + 
  ylab("Improvement") + 
  labs(title = "Forms") + 
  theme(axis.text=element_text(size=22),
        axis.title=element_text(size=30,face="bold"),
        title=element_text(size=25,face="bold"))
```



## Research Question 2: What, if any, area should we focus on for improvement?  E.g. are we better at teaching letters than we are at numbers?
```{r}
subset_scores$test_name <- factor(subset_scores$test_name,
    levels = c('preform','postform', 'prelet', 'postlet', 'prenumb', 'postnumb'),ordered = TRUE)


ggplot(subset_scores, aes(x=test_name, y=score, fill=which_test)) +
  geom_boxplot(show.legend=FALSE) +
  xlab("Which Test") + 
  ylab("Test Score") +
  labs(title = "Pre and Post Test Scores") +
  theme(axis.text=element_text(size=22),
        axis.title=element_text(size=30,face="bold"),
        axis.text.x = element_blank(),
        title=element_text(size=25,face="bold"))

ggplot(subset_scores, aes(x=which_test, y=score, fill=which_test)) +
  geom_boxplot(show.legend=FALSE) +
  xlab("Which Test") + 
  ylab("Test Score") +
  labs(title = "Average Test Scores") + 
  theme(axis.text=element_text(size=22),
        axis.title=element_text(size=30,face="bold"),
        title=element_text(size=25,face="bold"))
```

## Investigate some possible confounders

Make long format dataframe for confounders
```{r}

long_Data <- subset_scores |>
  left_join(Data, by = 'ID')

```


Sex vs avg test score
```{r}
ggplot(long_Data, aes(x = which_test, y = score, fill = sex)) + 
  geom_boxplot(show.legend=FALSE) +
  xlab("Which Test") +
  ylab("Test Score") + 
  labs(title = "Average Test Scores by Sex") +
  theme(axis.text=element_text(size=22),
        axis.title=element_text(size=30,face="bold"),
        title=element_text(size=25,face="bold"))
```

Site vs avg test score
```{r}
ggplot(long_Data, aes(x = which_test, y = score, fill = site)) + 
  geom_boxplot(show.legend=FALSE) +
  xlab("Which Test") +
  ylab("Test Score") + 
  labs(title = "Average Test Scores by Site") +
  theme(axis.text=element_text(size=22),
        axis.title=element_text(size=30,face="bold"),
        title=element_text(size=25,face="bold"))
```

Age vs avg test score
```{r}
ggplot(long_Data, aes(x = age, y = score, col = which_test)) + 
  geom_point() +
  xlab("Age") +
  ylab("Test Score") + 
  labs(title = "Average Test scores by Age", col = "Which Test") +
  theme(axis.text=element_text(size=22),
        axis.title=element_text(size=30,face="bold"),
        title=element_text(size=25,face="bold"),
        legend.text =element_text(size = 25, face = "bold"))
```

Setting vs avg test score
```{r}
ggplot(long_Data, aes(x = which_test, y = score, fill = setting)) + 
  geom_boxplot(show.legend=FALSE) +
  xlab("Which Test") +
  ylab("Test Score") + 
  labs(title = "Test scores by Setting") +
  theme(axis.text=element_text(size=22),
        axis.title=element_text(size=30,face="bold"),
        title=element_text(size=25,face="bold"),
        legend.text =element_text(size = 25, face = "bold"))
```
