---
title: "Sesame Street Report [Draft]"
author: 'Armelle, Sara, Ibrohim,'
date: 'February 05, 2024'
format: pdf
---

```{r}
#| echo: false
#| include: false
#| label: front-matter
#| warning: false
# clean up & set default chunk options
rm(list = ls())
knitr::opts_chunk$set(echo = FALSE)

# packages
library(tidyverse) # for example
library(mosaic)    # for example
library(ggformula) # for example
library(car)       # for example
library(tinytex)   # for example
```

## 1. Project Description

The project involves an observational study designed to assess the impact of
"Sesame Street" viewership on children's learning outcomes, particularly in
letters, numbers, and forms. The study is in the planning stage, aiming to
analyze existing data collected from children across five different sites,
including variables such as viewing frequency, setting, and encouragement to
watch, alongside pretest scores of vocabulary maturity (Peabody Picture
Vocabulary Test). The objective is to determine the program's effectiveness in
educational content delivery and identify areas for improvement. Results will
be used to enhance "Sesame Street's" educational focus and effectiveness, as
per the client's request for an upcoming board meeting presentation.

## 1.1 Research Questions

## 1.1 Research Questions

**Question 1:** How does viewership of "Sesame Street" influence children's learning outcomes, specifically in terms of letters, numbers, and forms knowledge, as measured by the percent increase in post-test scores compared to pre-test scores?

**Question 2:** What is the percent achievable gain (PAG) in learning outcomes for children who watch "Sesame Street," considering the maximum possible score improvement based on their initial knowledge level?

## 1.2 Variables

| Variable Type       | Variable Name          | Description                                                                                      | Measurement/Recording                                    |
|---------------------|------------------------|--------------------------------------------------------------------------------------------------|----------------------------------------------------------|
| Explanatory         | Viewing Frequency      | Frequency of "Sesame Street" viewership by children.                                             | Recorded as the number of episodes watched per week.     |
| Explanatory         | Viewing Setting        | Context in which "Sesame Street" is watched (e.g., at home, in school).                         | Categorical variable (Home, School, Other).              |
| Explanatory         | Encouragement to Watch | Level of encouragement from parents or teachers for children to watch "Sesame Street."           | Binary variable (Yes, No).                               |
| Response            | Percent Increase       | Difference in test scores before and after the exposure period, adjusted for scale differences.  | Calculated as percent_post_test - percent_pre_test.      |
| Response            | Percent Achievable Gain (PAG) | Improvement in score relative to the maximum possible improvement. | Calculated as score_diff/(max_score - pre_score). |

#  2. Exploratory Data Analysis (EDA)

The Exploratory Data Analysis (EDA) of the Sesame Street study integrates
comprehensive assessment and insights into children's learning improvements
attributed to the program. Through examining pre- and post-test outcomes over a
six-month interval among children aged 3-5, the study aims to quantify the
educational impact of Sesame Street on early learning metrics such as literacy
and numeracy. It meticulously considers variables like demographics, viewing
habits, and educational encouragement while addressing analytical challenges
like variance disparities, confounders, and multicollinearity. Initial findings
indicate variable program effectiveness, suggesting a correlation between
viewing frequency and educational advancement. The planned next steps involve
sophisticated statistical techniques to refine these insights, setting a robust
foundation for conclusive analysis and program enhancement strategies.

# 3. Statistical Analysis 

To effectively communicate the statistical analysis conducted in the Sesame
Street study, it's essential to balance technical accuracy with clarity for a
broader audience. The analysis began with verifying the assumptions of our
statistical models, such as linearity, independence, and normal distribution of
residuals, through diagnostic plots and tests, details of which are in the
appendix.

We interpreted the results in the context of the study's objectives. For
example, p-values were used to determine the statistical significance of the
relationship between Sesame Street viewership and learning outcomes. Lower
p-values indicated a stronger evidence against the null hypothesis, suggesting
that viewership does indeed impact learning outcomes.

The analysis, detailed enough for replication by another statistician, involved
comparing pre- and post-test scores, adjusting for confounding variables, and
employing ANOVA to understand the variance among different groups. This
approach allowed us to draw meaningful conclusions about the program's
effectiveness.

# 4. Recommendations

## Question 1

Our analysis indicates a positive correlation between Sesame Street viewership
and improved learning outcomes in children. Strategies to increase engagement
and content accessibility are recommended to leverage this educational impact.

## Question 2

The analysis of Percent Achievable Gain (PAG) suggests that personalized
content, based on the child's initial abilities, can significantly enhance
learning outcomes. We advise focusing on customized learning experiences.

# 5. Resources

- **Educational Technology Journals**: For cutting-edge research on educational
  media impacts.
- **Sesame Workshop's Research Portal**: Insights into educational content
  effectiveness.
- **Educator and Parent Online Communities**: For sharing strategies and
  experiences.

# 6. Additional Considerations

While our recommendations are based on current findings, variability in
individual learning responses and the long-term effects of media-based
education warrant further study. Concerns about demographic and contextual
factors suggest the need for tailored approaches. Future studies could benefit
from more sophisticated data analysis techniques and a broader consideration of
parental involvement.
  

# Technical Appendix  

*Detailed information and a copy of code and or software results.  Additional graphs and supporting figures may also be placed in the appendix.*


### R Script
```{r ref.label=c('front-matter', 'fig-example1','tbl-summarystats', 'fig-example2')}
#| echo: true
#| eval: false

# Reprinted code chunks used previously for analysis
```
