---
title: "Sesame_EDA"
author: "Armelle Duston"
date: "2024-01-31"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Libaries
```{r}
library(tidyverse)
library(ggplot2)
library(GGally)

```


## Import Data
```{r}
just_scores <- read.csv('test_scores.csv')
subset_scores <- subset(just_scores, 
                        grepl('let', test_name) | 
                        grepl('form', test_name) | 
                        grepl('numb', test_name))

subset_scores <- subset_scores |> 
  mutate(which_test = case_when(grepl('let', test_name) ~ 'letters',
                                grepl('form', test_name) ~ 'forms',
                                grepl('numb', test_name) ~ 'numbers'))

Data <- read.csv('clean_df.csv')

Data <-  Data |> 
  mutate(encour = case_when(encour == 0 ~ 'No',
                            encour == 1 ~ 'Yes'))

# make categorical variables categorical 
Data$site <-  as.character(Data$site)
Data$sex <-  as.character(Data$sex)
Data$viewcat <-  as.character(Data$viewcat)
Data$setting <-  as.character(Data$setting)


glimpse(Data)
```


## Research Question 1: Does our programming improve children’s knowledge of letters, numbers, and forms?

Comparing improvement (post-test - pre-test) based on viewcat (higher number = watched more)
```{r}

ggplot(Data, aes(x = viewcat, y = let_diff, fill = viewcat)) +
  geom_boxplot() +
  xlab("Viewing Category (high number = viewed more)") + 
  ylab("Letters Improvement (Posttest - Pretest)") + 
  labs(title = "Comparing Letters Outcomes by Viewing Category", 
       fill = "Viewing category")

ggplot(Data, aes(x = viewcat, y = numb_diff, fill = viewcat)) +
  geom_boxplot() +
  xlab("Viewing Category (high number = viewed more)") + 
  ylab("Numbers Improvement (Posttest - Pretest)") + 
  labs(title = "Comparing Numbers Outcomes by Viewing Category", 
       fill = "Viewing category")

ggplot(Data, aes(x = viewcat, y = form_diff, fill = viewcat)) +
  geom_boxplot() +
  xlab("Viewing Category (high number = viewed more)") + 
  ylab("Forms Improvement (Posttest - Pretest)") + 
  labs(title = "Comparing Forms Outcomes by Viewing Category", 
       fill = "Viewing category")

```

Comparing improvement (post-test - pre-test) based on encouragement (1 = was encouraged)
```{r}

ggplot(Data, aes(x = encour, y = let_diff, fill = encour)) +
  geom_boxplot() +
  xlab("Recieved Encouragement") +
  ylab("Letters Improvement (Posttest - Pretest)") + 
  labs(title = "Comparing Letters Outcomes by Viewing Encouragement",
       fill = "Recieved Encouragement")

ggplot(Data, aes(x = encour, y = numb_diff, fill = encour)) +
  geom_boxplot() +
  xlab("Recieved Encouragement") +
  ylab("Numbers Improvement (Posttest - Pretest)") + 
  labs(title = "Comparing Numbers Outcomes by Viewing Encouragement",
       fill = "Recieved Encouragement")

ggplot(Data, aes(x = encour, y = form_diff, fill = encour)) +
  geom_boxplot() +
  xlab("Recieved Encouragement") +
  ylab("Forms Improvement (Posttest - Pretest)") + 
  labs(title = "Comparing Forms Outcomes by Viewing Encouragement", 
       fill = "Recieved Encouragement")

```



## Research Question 2: What, if any, area should we focus on for improvement?  E.g. are we better at teaching letters than we are at numbers?
```{r}
subset_scores$test_name <- factor(subset_scores$test_name,
    levels = c('preform','postform', 'prelet', 'postlet', 'prenumb', 'postnumb'),ordered = TRUE)


ggplot(subset_scores, aes(x = test_name, y = score, fill = which_test)) +
  geom_boxplot() +
  xlab("Which Test") + 
  ylab("Test Score") +
  labs(title = "Comparing Pre- and Post- Test Scores for Forms, Letters, and Numbers",
       fill = "Which Test")

ggplot(subset_scores, aes(x = which_test, y = score, fill = which_test)) +
  geom_boxplot() +
  xlab("Which Test") + 
  ylab("Test Score") +
  labs(title = "Comparing Average Test Scores for Forms, Letters, and Numbers",
       fill = "Which Test")
```












