---
title: "Sesame_EDA"
author: "Armelle Duston"
date: "2024-01-31"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Libaries
```{r}
library(tidyverse)
library(ggplot2)
library(GGally)

```


## Import Data
```{r}
just_scores <- read.csv('test_scores.csv')
subset_scores <- subset(just_scores, 
                        grepl('let', test_name) | 
                        grepl('form', test_name) | 
                        grepl('numb', test_name))


wideData <- read.csv('clean_df_wide.csv')
longData <- read.csv('clean_df_long.csv')

# make categorical variables categorical 
wideData$site <-  as.character(wideData$site)
wideData$viewcat <-  as.character(wideData$viewcat)
wideData$setting <-  as.character(wideData$setting)

longData$site <-  as.character(longData$site)
longData$viewcat <-  as.character(longData$viewcat)
longData$setting <-  as.character(longData$setting)

glimpse(wideData)
```


## Research Question 1: Does our programming improve children’s knowledge of letters, numbers, and forms?

Comparing improvement (post-test - pre-test) based on viewcat (higher number = watched more)
```{r}

plot <- ggplot(wideData, aes(x = viewcat, y = letPAG, fill = viewcat)) +
  geom_boxplot(show.legend=FALSE) +
  xlab("View Time (higher value=more)") + 
  ylab("Percent Increase") + 
  labs(title = "Letters") 
  ggsave("boxLettersPAG.png", plot, width = 8, height = 6, dpi = 300)
plot

plot <- ggplot(wideData, aes(x = viewcat, y = numbPAG, fill = viewcat)) +
  geom_boxplot(show.legend=FALSE) +
  xlab("View Time") + 
  ylab("Percent Increase") + 
  labs(title = "Numbers") 
  ggsave("boxNumbersPAG.png", plot, width = 8, height = 6, dpi = 300)
plot

plot <- ggplot(wideData, aes(x = viewcat, y = formPAG, fill = viewcat)) +
  geom_boxplot(show.legend=FALSE) +
  xlab("View Time") + 
  ylab("Percent Increase") + 
  labs(title = "Forms") 
  ggsave("boxFormsPAG.png", plot, width = 8, height = 6, dpi = 300)
plot
```

Comparing improvement (post-test - pre-test) based on encouragement (1 = was encouraged)
```{r}

plot <- ggplot(wideData, aes(x = encour, y = letPAG, fill = encour)) +
  geom_boxplot(show.legend=FALSE) +
  xlab("Encouragement") + 
  ylab("Percent Increase") + 
  labs(title = "Letters") 
  ggsave("boxLettersPAG.png", plot, width = 8, height = 6, dpi = 300)
plot

plot <- ggplot(wideData, aes(x = encour, y = numbPAG, fill = encour)) +
  geom_boxplot(show.legend=FALSE) +
  xlab("Encouragement") + 
  ylab("Percent Increase") + 
  labs(title = "Numbers")
  ggsave("boxNumbersPAG.png", plot, width = 8, height = 6, dpi = 300)
plot

plot <- ggplot(wideData, aes(x = encour, y = formPAG, fill = encour)) +
  geom_boxplot(show.legend=FALSE) +
  xlab("Encouragement") + 
  ylab("Percent Increase") + 
  labs(title = "Forms") 
  ggsave("boxFormsPAG.png", plot, width = 8, height = 6, dpi = 300)
plot

```



## Research Question 2: What, if any, area should we focus on for improvement?  E.g. are we better at teaching letters than we are at numbers?
```{r}
# subset_scores$test_name <- factor(subset_scores$test_name,
#     levels = c('preform','postform', 'prelet', 'postlet', 'prenumb', 'postnumb'),ordered = TRUE)


# plot <- ggplot(subset_scores, aes(x=test_name, y=PAG, fill=which_test)) +
#   geom_boxplot(show.legend=FALSE) +
#   xlab("Which Test") + 
#   ylab("Test Score") +
#   labs(title = "Pre and Post Test Scores") +
#   theme(axis.text=element_text(size=22),
#         axis.title=element_text(size=30,face="bold"),
#         axis.text.x = element_blank(),
#         title=element_text(size=25,face="bold"))
#   ggsave("PrePostBox.png", plot, width = 8, height = 6, dpi = 300)

plot <- ggplot(longData, aes(x=which_test, y=PAG, fill=which_test)) +
  geom_boxplot(show.legend=FALSE) +
  xlab("Which Test") + 
  ylab("Percent Increase") +
  labs(title = "Percent Increase by Test") 
  ggsave("CategoriesBox.pngPAG", plot, width = 8, height = 6, dpi = 300)
plot
```


## Investigate some possible confounders

Sex vs avg test score
```{r}
plot <- ggplot(longData, aes(x = which_test, y = PAG, fill = sex)) + 
  geom_boxplot(show.legend=FALSE) +
  xlab("Which Test") +
  ylab("Percent Increase") + 
  labs(title = "Percent Increase by Sex") 
  ggsave("GenderBoxPAG.png", plot, width = 8, height = 6, dpi = 300)
plot
  

```

Site vs avg test score
```{r}
plot <- ggplot(longData, aes(x = which_test, y = PAG, fill = site)) + 
  geom_boxplot(show.legend=FALSE) +
  xlab("Which Test") +
  ylab("Percent Increase") + 
  labs(title = "Percent Increase by Site") 
  ggsave("SiteBoxPAG.png", plot, width = 8, height = 6, dpi = 300)
plot

```

Age vs avg test score
```{r}
plot <- ggplot(longData, aes(x = age, y = PAG, col = which_test)) + 
  geom_point() +
  xlab("Age") +
  ylab("Percent Increase") + 
  labs(title = "Percent Increase by Age", col = "Which Test") 
  ggsave("AgeScatterPAG.png", plot, width = 8, height = 6, dpi = 300)
plot 

```

Setting vs avg test score
```{r}

plot <- ggplot(longData, aes(x = which_test, y = PAG, fill = setting)) + 
  geom_boxplot(show.legend=FALSE) +
  xlab("Which Test") +
  ylab("Percent Increase") + 
  labs(title = "Percent Increase by Setting") 
  ggsave("SettingBoxPAG.png", plot, width = 8, height = 6, dpi = 300)
plot

```


## Do it all again but with pct diff


## Research Question 1: Does our programming improve children’s knowledge of letters, numbers, and forms?

Comparing improvement (post-test - pre-test) based on viewcat (higher number = watched more)
```{r}

plot <- ggplot(wideData, aes(x = viewcat, y = let_diff_pct, fill = viewcat)) +
  geom_boxplot(show.legend=FALSE) +
  xlab("View Time (higher value=more)") + 
  ylab("Percent Increase") + 
  labs(title = "Letters") 
  ggsave("boxLettersPCT.png", plot, width = 8, height = 6, dpi = 300)
plot

plot <- ggplot(wideData, aes(x = viewcat, y = numb_diff_pct, fill = viewcat)) +
  geom_boxplot(show.legend=FALSE) +
  xlab("View Time") + 
  ylab("Percent Increase") + 
  labs(title = "Numbers") 
  ggsave("boxNumbersPCT.png", plot, width = 8, height = 6, dpi = 300)
plot

plot <- ggplot(wideData, aes(x = viewcat, y = form_diff_pct, fill = viewcat)) +
  geom_boxplot(show.legend=FALSE) +
  xlab("View Time") + 
  ylab("Percent Increase") + 
  labs(title = "Forms") 
  ggsave("boxFormsPCT.png", plot, width = 8, height = 6, dpi = 300)
plot
```

Comparing improvement (post-test - pre-test) based on encouragement (1 = was encouraged)
```{r}

plot <- ggplot(wideData, aes(x = encour, y = let_diff_pct, fill = encour)) +
  geom_boxplot(show.legend=FALSE) +
  xlab("Encouragement") + 
  ylab("Percent Increase") + 
  labs(title = "Letters") 
  ggsave("boxLettersPCT.png", plot, width = 8, height = 6, dpi = 300)
plot

plot <- ggplot(wideData, aes(x = encour, y = numb_diff_pct, fill = encour)) +
  geom_boxplot(show.legend=FALSE) +
  xlab("Encouragement") + 
  ylab("Percent Increase") + 
  labs(title = "Numbers")
  ggsave("boxNumbersPCT.png", plot, width = 8, height = 6, dpi = 300)
plot

plot <- ggplot(wideData, aes(x = encour, y = form_diff_pct, fill = encour)) +
  geom_boxplot(show.legend=FALSE) +
  xlab("Encouragement") + 
  ylab("Percent Increase") + 
  labs(title = "Forms") 
  ggsave("boxFormsPCT.png", plot, width = 8, height = 6, dpi = 300)
plot

```



## Research Question 2: What, if any, area should we focus on for improvement?  E.g. are we better at teaching letters than we are at numbers?
```{r}
# subset_scores$test_name <- factor(subset_scores$test_name,
#     levels = c('preform','postform', 'prelet', 'postlet', 'prenumb', 'postnumb'),ordered = TRUE)


# plot <- ggplot(subset_scores, aes(x=test_name, y=diff_pct, fill=which_test)) +
#   geom_boxplot(show.legend=FALSE) +
#   xlab("Which Test") +
#   ylab("Test Score") +
#   labs(title = "Pre and Post Test Scores") +
#   ggsave("PrePostBoxPCT.png", plot, width = 8, height = 6, dpi = 300)

plot <- ggplot(longData, aes(x=which_test, y=diff_pct, fill=which_test)) +
  geom_boxplot(show.legend=FALSE) +
  xlab("Which Test") + 
  ylab("Percent Increase") +
  labs(title = "Percent Increase by Test") 
  ggsave("CategoriesBoxPCT.png", plot, width = 8, height = 6, dpi = 300)
plot
```


## Investigate some possible confounders



Sex vs avg test score
```{r}
plot <- ggplot(longData, aes(x = which_test, y = diff_pct, fill = sex)) + 
  geom_boxplot(show.legend=FALSE) +
  xlab("Which Test") +
  ylab("Percent Increase") + 
  labs(title = "Percent Increase by Sex") 
  ggsave("GenderBoxPCT.png", plot, width = 8, height = 6, dpi = 300)
plot
  

```

Site vs avg test score
```{r}
plot <- ggplot(longData, aes(x = which_test, y = diff_pct, fill = site)) + 
  geom_boxplot(show.legend=FALSE) +
  xlab("Which Test") +
  ylab("Percent Increase") + 
  labs(title = "Percent Increase by Site") 
  ggsave("SiteBoxPCT.png", plot, width = 8, height = 6, dpi = 300)
plot

```

Age vs avg test score
```{r}
plot <- ggplot(longData, aes(x = age, y = diff_pct, col = which_test)) + 
  geom_point() +
  xlab("Age") +
  ylab("Percent Increase") + 
  labs(title = "Percent Increase by Age", col = "Which Test") 
  ggsave("AgeScatterPCT.png", plot, width = 8, height = 6, dpi = 300)
plot 

```

Setting vs avg test score
```{r}

plot <- ggplot(longData, aes(x = which_test, y = diff_pct, fill = setting)) + 
  geom_boxplot(show.legend=FALSE) +
  xlab("Which Test") +
  ylab("Percent Increase") + 
  labs(title = "Percent Increase by Setting") 
  ggsave("SettingBoxPCT.png", plot, width = 8, height = 6, dpi = 300)
plot

```







